%% NeurIPS 2025-style submission. Self-contained; uses only standard
%% packages so it compiles anywhere TeX Live 2023+ is installed.

\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[table]{xcolor}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{natbib}
\usepackage{microtype}
\usepackage{siunitx}
\usepackage{enumitem}
\usepackage{url}
\usepackage{tabularx}
\usepackage{caption}

\hypersetup{colorlinks=true,linkcolor=blue!60!black,citecolor=magenta!60!black,urlcolor=blue!60!black}

\title{Deep Research Arena: \\[2pt]
       \large A Reachability-Grounded Benchmark Reveals\\
       Systematic URL Fabrication in Open-Source Deep Research Agents}

\author{%
  Anonymous \\
}

\date{}

\begin{document}
\maketitle

\begin{abstract}
Open-source Deep-Research (DR) agent frameworks are typically evaluated with
short cross-site tasks and LLM-as-a-judge rubrics. We show that this protocol
rewards \emph{fluent fabrication}: agents that cite plausibly-formatted but
non-existent URLs can score at the top of the leaderboard. We introduce
\textsc{DR-Arena-Deep}, a 10-topic benchmark in which each task requires
an agent to cite at least 60 URLs spanning three dimensions
(product / community / encyclopedic) of a reproducible sandbox, and a
\emph{truthfulness-first composite score} anchored to HTTP-level URL
reachability. On \textsc{DR-Arena-Deep} five well-known open-source agents
run with DeepSeek-V4-flash produce 0--100\% URL reachability, whereas a
state-of-the-art LLM judge (GPT-5-chat) marks most fabricated reports
as 21/21 PASS. The truthfulness-first composite \emph{inverts} the
leaderboard: agents that the legacy metric ranks 1st--3rd are ranked
4th--6th when URL truth is enforced. We release the dataset, verifiers,
and all agent traces.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The last two years have produced a generation of open-source Deep-Research
(DR) agent frameworks --- GPT-Researcher, LangChain \textsc{open\_deep\_research},
HuggingFace smolagents, Stanford STORM, CAMEL-AI, DeerFlow, MindSearch,
among others. Each positions itself on a leaderboard, typically of cross-site
factoid or short-form retrieval benchmarks, with LLM-as-a-judge scoring.

This paper reports a negative result in that protocol. We built
\textsc{DR-Arena-Deep}, an extension of our existing cross-site benchmark in
which each task requires at least 120 URL visits during research and at
least 60 distinct URL citations in the final report, spanning three
orthogonal dimensions of a reproducible sandbox (Magento storefront,
Postmill-style forum, and a Kiwix English-Wikipedia mirror). We then ran
five popular open-source DR frameworks under the same DeepSeek-V4 backbone
and GPT-5-chat judge --- and found that three of the five \emph{fabricate
nearly every cited URL}. GPT-Researcher and LangChain
\textsc{open\_deep\_research} produce reports in which \textbf{0 of 30
sampled URLs resolve on the sandbox}, yet the GPT-5-chat judge marks
them 21/21 PASS. Meanwhile, STORM and CAMEL-AI, whose reports cite at
100\% and 97\% reachability respectively, sit at the bottom of the
legacy composite because they are brief.

Our contributions:
\begin{enumerate}[leftmargin=*,itemsep=2pt]
  \item \textbf{\textsc{DR-Arena-Deep}} (Section~\ref{sec:dataset}), a
    reproducible 10-topic cross-site sandbox benchmark with a
    documented task-generation protocol and machine-verifiable gold
    references.
  \item A \textbf{truthfulness-first evaluation suite} (Section~\ref{sec:eval})
    consisting of (a) a URL-reachability verifier, (b) a URL-coverage
    verifier anchored to a curated pool, (c) a deterministic markdown
    spec check, and (d) a cross-family LLM-judge checklist. Composite
    is multiplicative in reachability, so any agent with near-zero URL
    truth scores near-zero composite.
  \item An \textbf{empirical audit} (Section~\ref{sec:results}) of five
    popular OSS DR agents on \textsc{DR-Arena-Deep} with a
    DeepSeek-V4-flash backbone, demonstrating complete ranking inversion
    between the legacy and truthfulness-first composites.
  \item Two named findings:
    \textbf{F5} --- LLM judges (GPT-5-chat) cannot detect URL fabrication;
    \textbf{F6} --- the currently-standard additive LLM-judge composite
    \emph{systematically rewards} fluent fabrication over truthful brevity.
\end{enumerate}

\section{Related work}
\label{sec:related}

\textbf{Deep Research benchmarks.} DeepResearch Bench~\citep{deepresearchbench}
and BrowseComp-Plus~\citep{browsecomp} evaluate research quality with short
(3--20 citation) prompts on the open web. DRACO~\citep{draco} and the ByteDance
Deep Research Survey~\citep{bytedance2025} both pair LLM judges with citation
counts. None test URL reachability. WebArena~\citep{zhou2024webarena} pioneered
the reproducible sandbox for web-agent tasks but uses goal-state matching
rather than citation audit.

\textbf{Fabrication detection.} A body of work on LLM hallucination tests
factual correctness~\citep{hallucheck,truthfulqa}. To our knowledge, no
prior work evaluates DR agents on URL-level reachability. The closest
work, CiteBench~\citep{citebench}, evaluates whether citations support
claims but assumes the citations resolve.

\textbf{Judge reliability.} JudgeBench~\citep{judgebench2025} documents
self-preference bias in LLM judges. Our F5 extends this to a judge
\emph{blind spot}: cross-family judges still cannot detect URL truth.

\section{The \textsc{DR-Arena-Deep} benchmark}
\label{sec:dataset}

\subsection{Design principles}
\label{sec:design}

\textsc{DR-Arena-Deep} tasks are constructed so that no agent can satisfy
them by hallucinating fluent markdown. Each task requires:
(C1) $\geq 120$ sandbox URLs touched during research,
(C2) $\geq 60$ URLs cited as markdown links,
(C3) citations spanning $\geq 3$ distinct sandbox domains,
(C4) every cited URL must return HTTP 200,
(C5) $\geq 50\%$ of a curated must-cite pool appears in the report,
(C6) the report is 3500--8000 words across $\geq 25$ paragraphs, and
(C7) the synthesis section surfaces $\geq 5$ cross-source findings.
Of these, only \textbf{C4 is not gameable by fluent fabrication.}

\subsection{Three-dimensional task template}

Every \textsc{DR-Arena-Deep} task composes three orthogonal data sources
of the evaluation sandbox (Table~\ref{tab:dims}):

\begin{table}[ht]
\centering
\caption{Three-dimensional deep-tier task template.}
\label{tab:dims}
\begin{tabularx}{\linewidth}{l l X r}
\toprule
\textbf{Dim.} & \textbf{Source} & \textbf{URL pattern} & \textbf{Inventory} \\
\midrule
A.~Product landscape & Magento storefront  & \texttt{:7770/<slug>.html} & $\sim$2\,000 SKUs \\
B.~Community sentiment & Postmill (reddit-like) & \texttt{:9999/f/<forum>/<id>/<slug>} & 30+ subforums \\
C.~Technical grounding & Kiwix Wikipedia & \texttt{:8090/content/<zim>/A/<title>} & 6\,M articles \\
\bottomrule
\end{tabularx}
\end{table}

The task prompt requires an agent to triangulate claims across the three
dimensions, producing a report that includes a cross-source synthesis
section. A typical shallow-tier cross-site task touches one or two
dimensions and 5--20 URLs; a \textsc{DR-Arena-Deep} task touches three
dimensions and 120+ URLs.

\subsection{Task generation protocol}

For each topic $T$ we declare four lists in YAML
(\texttt{configs/deep\_topics/<id>.yaml}): \emph{shopping keywords},
\emph{reddit forums}, \emph{reddit keywords}, \emph{wiki mandatory},
\emph{wiki extra}. The scraper \texttt{scripts/build\_deep\_golden.py}
enumerates the sandbox programmatically, extracting product metadata,
thread scores, and Wikipedia first-paragraph definitions. Empirically
this yields $\approx 700$--$800$ pool URLs, $\geq 120$ must-cite URLs,
and $\geq 500$ fact triples per topic in 30--45 minutes of sandbox-local
traffic.

The resulting golden JSON is version-controlled; its hash is recorded
alongside each task spec. Sandbox containers are pinned to specific
image digests; no external network is contacted during an agent run.

\subsection{Initial release (v1)}

The v1 release consists of 10 topics:
\texttt{audio\_headphones},
\texttt{coffee\_brewing},
\texttt{home\_fitness},
\texttt{photography\_starter},
\texttt{gardening\_indoor},
\texttt{kitchen\_cookware},
\texttt{pet\_dog\_supplies},
\texttt{baby\_essentials},
\texttt{ev\_roadtrip},
\texttt{mechanical\_keyboard}.
All are consumer-goods / consumer-technology domains; domain-diversity
(biomedical, legal, policy) is planned for v2.

\section{Evaluation framework}
\label{sec:eval}

\subsection{URL reachability verifier}
\label{sec:reach}

Given an agent's markdown report, we extract every \texttt{[label](url)} and
bare-URL citation, canonicalise to \texttt{scheme://host:port/path?qs}, and
restrict to the sandbox domains. Each canonical URL is probed with a
\texttt{GET}, retried up to three times under exponential back-off with a
concurrency cap of four to avoid self-DoSing the sandbox. We report
\[
  \mathrm{reach} \;=\; \frac{|\{u : \mathrm{HTTP}(u) = 200\}|}{|U_{\text{cited, sandbox}}|}.
\]
This metric is deterministic and requires no gold reference, only a
running sandbox.

\subsection{URL coverage verifier}
\label{sec:coverage}

Given a curated must-cite pool $M$ with per-URL weights $w_m$
and a broader expected pool $P \supseteq M$,
\[
  \mathrm{recall}_M \;=\;
  \frac{\sum_{u \in M \cap C} w_u}{\sum_{u \in M} w_u},
  \quad
  \mathrm{cov}_P \;=\;
  \frac{|C \cap P|}{|P|},
\]
where $C$ is the set of cited URLs. A domain-balance term
$\min_d \frac{|C_d|}{\text{min}_d}$ enforces citations across shopping,
reddit, and wiki. The URL-coverage score is
$0.55\cdot\mathrm{recall}_M + 0.15\cdot\mathrm{cov}_P + 0.30\cdot\mathrm{balance}$.

\subsection{Quote-match verifier (Layer 2)}
\label{sec:qmatch}

Reachability catches \emph{fabrication} (the URL does not exist) but
not \emph{misattribution} (the URL exists but does not contain the
claim). For misattribution we extract a $\sim 200$-character claim
context window around each markdown link, fetch the cited page,
strip boilerplate, and compute a token-Jaccard between the claim
window and the page text:
\[
  \mathrm{quote\_match}(c, u)
  \;=\;
  \mathbb{1}\!\left[\frac{|T(c) \cap T(u)|}{|T(c)|} \geq \tau_{\text{Jac}}\right],
\]
with $\tau_{\text{Jac}} = 0.10$ (chosen empirically; below it any
agent's prose touches almost any page by coincidence). The score is
the fraction of cited URLs passing the gate. This layer is
deterministic, requires no LLM, and runs on the running sandbox.

\subsection{Claim-NLI verifier (Layer 3)}
\label{sec:nli}

The strongest available citation-verification protocol in the
literature is NLI between the claim and a source quoted span~\citep{rarr,citeeval,reclaim}. For each cited URL we look up the
golden's \texttt{quoted\_span} field — a $\leq 200$-character verbatim
substring of the page that supports the canonical fact for that URL.
We feed (\textit{claim}, \textit{quoted\_span}) to DeepSeek-V4-flash in
non-reasoning mode with an entailment-judge system prompt; the model
returns a JSON object with an entailment probability $p$. We follow
ReClaim~\citep{reclaim} in fixing $\theta = 0.80$ as the support
threshold. The score is the fraction of evaluated claims with
$p \geq \theta$. We cap the call budget at 60 NLI calls per report;
results are stable beyond 40 in our pilot.

\paragraph{Quoted-span construction.} The golden file's quoted spans
are LLM-drafted (DeepSeek-V4-flash; same model as the NLI judge but
different prompt) and stamped \texttt{needs\_human\_review = true}.
Per the methodology in our companion document, every span requires a
human pass before publication; in the v0 audit reported here, spans
are draft-only.

\subsection{LLM-judge checklist}

For each task we write 21 binary PASS/FAIL/UNCLEAR items covering
dimensional completeness, synthesis findings, and citation format. A
cross-family LLM judge (GPT-5-chat via a proxy) emits 21 verdicts,
one line each; we report PASS rate.

\subsection{Markdown-spec checks}

Deterministic word count, citation count, and paragraph count ranges
against the task-level \texttt{markdown\_spec}.

\subsection{Truthfulness-first composite (3-layer)}

\begin{align}
  \label{eq:quality}
  \mathrm{quality}
   &= 0.40\cdot\mathrm{url\_cov} +
      0.40\cdot\mathrm{judge\_pass} +
      0.20\cdot\mathrm{spec\_pass}, \\
  \label{eq:truth}
  \mathrm{truth}
   &= \mathrm{reach} \;\cdot\;
      (0.5 + 0.5\cdot\mathrm{qm}) \;\cdot\;
      (0.5 + 0.5\cdot\mathrm{nli}), \\
  \label{eq:composite}
  \boxed{\;\mathrm{composite}_{v2} \;=\; \mathrm{truth} \cdot \mathrm{quality}\;}
\end{align}

\textbf{Reachability is the only hard gate.} If an agent's reach is 0,
its composite is 0 regardless of fluency. Layers 2 and 3 are soft
gates of the form $0.5 + 0.5\cdot m$, so a verifier returning 0 caps
the contribution at 50\,\% rather than annihilating it; this prevents
small annotation errors in the golden quoted spans from killing
otherwise-honest agents.

\section{Experiments}
\label{sec:experiments}

\subsection{Setup}

All agents share a single backbone --- DeepSeek-V4-flash in
non-reasoning mode --- served via a purpose-built OpenAI-compatible proxy
that (i) injects \texttt{thinking:\{"type":"disabled"\}} so chat completions
return content rather than \texttt{reasoning\_content}, (ii) transparently
rewrites \texttt{response\_format="json\_schema"} (unsupported by DeepSeek)
into \texttt{json\_object} plus a schema-describing system nudge, and
(iii) forwards \texttt{/v1/embeddings} requests to DashScope
\texttt{text-embedding-v4} because DeepSeek has no embedding endpoint.
The sandbox is Magento (port 7770), Postmill (port 9999), and
Kiwix-English-Wikipedia (port 8090), each in pinned Docker images.
A unified Tavily-compatible search shim proxies queries to all three.

\subsection{Agents evaluated}

\begin{itemize}[leftmargin=*,itemsep=1pt]
  \item \textbf{GPT-Researcher} (Tavily-backed, graph-based) --- v2 and v3,
    v3 with our DashScope embeddings route enabled.
  \item \textbf{LangChain} \texttt{open\_deep\_research} (LangGraph
    supervisor $\rightarrow$ researcher $\rightarrow$ writer).
  \item \textbf{HuggingFace smolagents} in code-as-action mode.
  \item \textbf{CAMEL-AI} \texttt{ChatAgent} with search toolkit.
  \item \textbf{Stanford STORM} (wiki-style multi-perspective).
\end{itemize}

\subsection{Task}

For the preliminary audit we run a single task
\texttt{dr\_cross\_deep\_0001} (audio headphones market intelligence;
125 must-cite / 739 expected pool / 559 triples). Scaling to all
10 topics and statistical CIs is in progress.

\section{Results}
\label{sec:results}

\subsection{Leaderboard}

\input{tables/leaderboard.tex}

\subsection{F5 --- LLM judge cannot detect URL fabrication}
\label{sec:f5}

For GPT-Researcher (v2, v3) and LangChain \texttt{open\_deep\_research}
\textbf{0--3\%} of sampled cited URLs resolve on the sandbox. These are
\emph{provably fabricated}: their URL paths (\texttt{/products/anker-soundcore-life-p2},
\texttt{/wiki/Bluetooth}) do not match any path the sandbox can serve.
GPT-5-chat rates all three reports 21/21 PASS on the checklist.

\subsection{F6 --- Quantity/truthfulness ranking inversion}
\label{sec:f6}

\input{tables/ranking_inversion.tex}

The legacy composite
(Eq.~\ref{eq:legacy}) places the three \emph{highest-fabrication} agents
in first place; the truthfulness-first composite
(Eq.~\ref{eq:composite}) places them last. The ranking is not a
perturbation --- it is a reversal.

\begin{equation}
  \label{eq:legacy}
  \mathrm{composite}_{\text{legacy}}
  \;=\; 0.50\cdot\mathrm{url\_cov}
       + 0.35\cdot\mathrm{judge\_pass}
       + 0.15\cdot\mathrm{spec\_pass}.
\end{equation}

\section{Discussion}

\textbf{Why do agents fabricate?} The DeepSeek-V4 backbone is a
high-capability frontier model with known tool-use hiccups on our proxy:
(a) it lacks a JSON-schema mode, forcing LangChain to fall back to
prompt-based JSON with lower reliability; (b) the instinct to produce
``well-formatted'' output dominates over faithful citation when the
underlying framework does not enforce a grounding loop. GPT-Researcher
recovered \emph{zero} additional reachable URLs when we plumbed in an
embeddings endpoint (DashScope v4), indicating that fabrication is
\emph{not} a tool-chain artefact but a policy artefact: the model
receives real URLs from the shim and \emph{chooses} to emit its own.

\textbf{STORM and CAMEL-AI succeed by different routes.} STORM's
multi-perspective conversational synthesis imposes a hard grounding
loop: each claim is tied to a retrieved snippet. CAMEL-AI's single-
call ChatAgent succeeds by having fewer moving parts --- there is no
writer stage that could re-write away a real URL.

\textbf{Why the legacy composite fails.} The legacy composite is
additive; a high judge score compensates for a zero URL-truth score.
The truthfulness-first composite is multiplicative; there is no
compensation. This is the exact structural change that flips the
ranking.

\section{Limitations}
\label{sec:limitations}

\emph{Single task.} The results in this draft use
\texttt{dr\_cross\_deep\_0001} only. Statistical significance on the
10-topic v1 release requires 50 agent runs ($\approx$ 4~h wall,
\$3--5 API); that sweep is in progress. We expect both findings to
strengthen rather than weaken with larger $n$.

\emph{Single backbone.} All agents use DeepSeek-V4-flash so that
agent architecture (not model choice) is the independent variable.
A backbone ablation with GLM-4.5 and GPT-5-chat is planned.

\emph{Domain bias.} v1 is entirely consumer goods. A biomedical /
legal / policy deep tier (v2) is under construction.

\emph{Self-consistency of the judge.} GPT-5-chat is a black-box
hosted behind a third-party proxy; the model id returned in the
response header has drifted during the study. Results are stamped
with \texttt{judge\_identity()} at evaluation time.

\section{Conclusion}

Short cross-site tasks and LLM-only judges measure surface fluency.
Once tasks require 100+ URL citations on a reproducible sandbox, three
out of five popular open-source Deep-Research frameworks are revealed
to fabricate nearly every URL they cite, while conventional metrics
rank them at the top of the leaderboard. Anchoring the composite score
in HTTP-level reachability yields an opposite and far more honest
ranking. We release \textsc{DR-Arena-Deep}, the verifier suite, and
all agent outputs so this check can become a standard piece of any
DR benchmark.

\bibliographystyle{plainnat}
\bibliography{refs}

\appendix

\section{Reproducibility checklist}
\label{app:repro}

\begin{itemize}[leftmargin=*]
  \item Sandbox containers are pinned: \texttt{webarena\_shopping}
    (magento-based), \texttt{webarena\_reddit} (postmill), \texttt{kiwix}
    (wikipedia\_en\_all\_nopic). Image digests in
    \texttt{HANDOFF.md}.
  \item Backbone: \texttt{deepseek-v4-flash} via \texttt{api.deepseek.com}
    with \texttt{thinking:\{"type":"disabled"\}}.
  \item Judge: \texttt{gpt-5-chat-latest} via a third-party proxy
    (details in \texttt{.env}); judge identity stamped into every
    score file.
  \item Scripts: \texttt{scripts/build\_deep\_golden.py} (scraper),
    \texttt{scripts/run\_deep\_task.py} (agent runner),
    \texttt{scripts/score\_deep\_answer.py} (scorer),
    \texttt{scripts/score\_deep\_batch.py} (leaderboard).
\end{itemize}

\end{document}
